\mytitle

### 2021-12-15 paper.tex
improve formulation
refine abstact and title

\newcommand{\mytitle}{An algorithm for the calculation of the real dilogarithm with instruction-level parallelism}
\newcommand{\myauthor}{Alexander Voigt}
\newcommand{\Li}{\operatorname{Li}_2}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}

@@

An algorithm for the computation of the real dilogarithm is presented,
which allows for the use of instruction\-/level parallelism when
implemented on appropriate CPUs.

@@

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures with
    different compilers.}
  \begin{tabular}{lcc}
    \toprule
    \multirow{2}{*}{Implementation} & i7-4700MQ & i7-5600U \\
    & \texttt{g++} 10.2.1 & \texttt{g++} 9.3.0 \\
    \midrule
    this paper & $0.031$ & $0.026$ \\
    Morris (DILOG 0011) \cite{morris} & $0.035$ & $0.026$ \\
    Cephes \cite{cephes} & $0.044$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.054$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.079$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.125$ & $0.071$ \\
    GSL \cite{gsl} & $0.330$ & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper is fastest
among the ones shown in the table.  The next-to fastest implementation
is provided by Cephes, which is only $\sim 20\%$ slower.  The short
runtime of Cephes stems from the fact that Cephes also uses a rational
function approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.

>>>>

\newcommand{\mytitle}{Comparisons of methods for the calculation of the real dilogarithm regarding instruction-level parallelism}
\newcommand{\myauthor}{Alexander Voigt}
\newcommand{\Li}{\operatorname{Li}_2}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}

@@

We compare different methods for the computation of the real
dilogarithm regarding their ability for using instruction\-/level
parallelism when executed on appropriate CPUs.  As a result we present
an instruction\-/level\-/aware method and compare it to existing
implementations.

@@

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures with
    different compilers.}
  \begin{tabular}{lcc}
    \toprule
    \multirow{2}{*}{Implementation} & i7-4700MQ & i7-5600U \\
    & \texttt{g++} 10.2.1 & \texttt{g++} 9.3.0 \\
    \midrule
    this paper & $0.031$ & $0.026$ \\
    Morris (DILOG 0011) \cite{morris} & $0.035$ & $0.026$ \\
    Cephes \cite{cephes} & $0.044$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.054$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.079$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.125$ & $0.071$ \\
    GSL \cite{gsl} & $0.330$ & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper and the
approach presented in \cite{morris} are fastest among the ones shown
in the table.  The next-to fastest implementation is provided by
Cephes, which is only $\sim 20\%$ slower.  The short runtime of Cephes
stems from the fact that Cephes also uses a rational function
approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.
